<<<<<<< HEAD
---
title: "P8105-HW2-lj2607"
author: "Li Jiang"
date: "2024-10-01"
output: github_document
---

#Problem 1

```{r}
#load the required packages
library(tidyverse)
library(readxl)
```

```{r}
#clean the data and retain the variable as required 
trans_ent = 
  read_csv(
    "Datasets/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) |> 
  janitor::clean_names() |> 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

#visually check the dataset
view(trans_ent)
```

```{r}
#check the size of dataset

nrow(trans_ent)
ncol(trans_ent)
```

Description of the dataset nyc_transit_clean:

I currently cleaned the original dataset and retained line, station_name, station_latitude, station_longitude, route, entry, exit_only, vending, entrance_type, ada. I also converted entry to logical variable. The sesulting dataset has `r ncol(trans_ent)` columns and `r nrow(trans_ent)` rows. The dimenison of the resulting data set is : `r dim(trans_ent)`. The dataset contain variables : `r names(trans_ent)`. As it stands, these data are not "tidy": route number should be a variable, as should route.

```{r}
#Number of distinct station
trans_ent |> 
  select(station_name, line) |> 
  distinct()
```
There are 465 distinct station.

```{r}
#Number of distinct station that are ADA compliant
trans_ent |> 
  filter(ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```
There are 84 ADA compliant station.

```{r}
#compute the proportion of station entrances / exits without vending allow entrance
trans_ent |> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```
The proportion of station entrances / exits without vending allow entrance is 0.3770492


```{r}
#identify stations that serve the A train
trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()
```

There are 60 distinct stations servcing the A train.

```{r}
#assess how many of these are ADA compliant
trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A", ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```

17 of these are ADA compliant


#Problem 2

```{r}
# Import Mr.Trash Wheel dataset, specifying the sheet and range

mr_trash_data <- read_excel("Datasets/202409 Trash Wheel Collection Data.xlsx",
                             sheet = 'Mr. Trash Wheel',
                             range = "A2:N653",
                             na = c("NA", "", ".")) %>% 
                   janitor::clean_names() %>% # use reasonable variable names
                   drop_na(dumpster) %>%      # drop the rows without specific dumpster                                                           information
                   mutate(sports_balls = as.integer(round(sports_balls,digits = 0)),
                                              # convert sport ball to integer variable
                          trash_wheel = "mr_trash", # creating additional variable recording                                                            trash_wheel           
                          year = as.numeric(year)) # convert year to numerical variable
                                             

# visually inspect dataset
view(mr_trash_data)
```

```{r}
#Import Professor.Trash Wheel dataset, specifying the sheet and range

prof_trash_data <- read_excel("Datasets/202409 Trash Wheel Collection Data.xlsx",
                              sheet = 'Professor Trash Wheel',
                              range = "A2:M120",
                              na = c("NA", "", ".")) %>% 
                   janitor::clean_names() %>% # use reasonable variable names
                   drop_na(dumpster) %>%      # drop the rows without specific dumpster                                                           information
                   mutate(trash_wheel = "prof_trash") 
                                              # creating additional variable recording trash_wheel        

# visually inspect dataset
view(prof_trash_data)

#Total weight of trash collected by professor trash wheel

sum(prof_trash_data$weight_tons, na.rm = TRUE)
```

```{r}
#Import Gwynnda Trash Wheel dataset, specifying the sheet and range

gwynnda_trash_data <- read_excel("Datasets/202409 Trash Wheel Collection Data.xlsx",
                              sheet = 'Gwynnda Trash Wheel',
                              range = "A2:L265",
                              na = c("NA", "", ".")) %>% 
                   janitor::clean_names() %>% # use reasonable variable names
                   drop_na(dumpster) %>%      # drop the rows without specific dumpster                                                           information
                   mutate(trash_wheel = "gwynnda_trash") 
                                              # creating additional variable recording trash_wheel        

# visually inspect dataset
view(gwynnda_trash_data)

# total number of cigarette butts collected by Gwynnda in June of 2022
total <- gwynnda_trash_data %>%
         filter(month == "June", year == "2022") %>%
         summarise(total_cigarette_butts = sum(cigarette_butts, na.rm = TRUE))
total
```

```{r}
#Combine three sheets to produce a single tidy sheet
trash_wheel_data = 
  bind_rows(mr_trash_data, prof_trash_data, gwynnda_trash_data) %>% 
  janitor::clean_names()

# visually inspect dataset
view(trash_wheel_data)
```

Trash_wheel_data is composed of three datasets: mr_trash_data, prof_trash_data, gwynnda_trash_data. There are `r nrow(trash_wheel_data)` rows and `r ncol(trash_wheel_data)` in datase and key variable include `r names(trash_wheel_data)`. The number of observation in datset is `r nrow(trash_wheel_data)`. The total weight of trash collected by professor trash wheel is 246.74 tons. The total number of cigarette butts collected by Gwynnda in June of 2022 is 18120.


##problem 3

```{r}
#Import baker dataset

bakers <- read_csv("Datasets/bakers.csv",
                   na = c("NA", "", ".")) %>%
          janitor::clean_names() %>%
          separate(baker_name, into = c("baker", "last_name"), sep = " ")

#visually inspect the dataset
view(bakers)
```

```{r}
#Import bakes dataset

bakes <- read_csv("Datasets/bakes.csv",
                   na = c("NA", "", ".")) %>%
         janitor::clean_names()

#visually inspect the dataset
view(bakes)
```

```{r}
#Import results dataset

results <- read_csv("Datasets/results.csv",
                   na = c("NA", "", "."),skip=2) %>%
          janitor::clean_names()%>%
          mutate(series = as.numeric(series),
                 episode = as.numeric(episode))

#visually inspect the dataset
view(results)
```

```{r}
#use anti-join to detect differences--missing value in bake
missing_bake <- anti_join(bakers,bakes)
view(missing_baker)

#use anti-join to detect differences--missing value in results
missing_results <- anti_join(bakers,results, by = "baker")
view(missing_results)
```

```{r}
#Merge three datasets
results_bakes <- left_join(bakes, results, by = c("baker", "series", "episode"))

fianl_merge <- left_join(results_bakes, bakers, by = c("baker", "series"))%>%
               select(series, episode, baker, baker_last_name, baker_age, baker_occupation, 
                      hometown)
view(fianl_merge)
```

I imported three datasets: bakers, bakes, and results, assigned corresponding NA values, and cleaned the variable names. I split the full names into two first name and last names and convert episode and series to numeric variable. I was able to identify the missing result as jo wheatley.

Next, I used left_join to merge all three datasets into one dataset. This combined dataset contains details about the bakers' backgrounds (age, hometown, etc.), signature bakes, showstoppers, and their results (whether they progressed or were eliminated). Since the bakes dataset is incomplete, I retain the following variable:series, episode, baker, baker_last_name, baker_age, baker_occupation, hometown

```{r}
#create a reader-friendly table showing the winner in season 5 through 10
winners <-
  results %>% 
  filter(series <= 10, series >=5) %>% 
  filter(result == c("WINNER", "STAR BAKER"))%>% 
  select(series, episode, baker)
view(winners)

winners_table <-winners %>%
                pivot_wider(
                names_from = series,
                values_from = baker)
knitr::kable(winners_table)
```
I created a reader-friendly table based on `results` dataset and presented winners across seasons 5 to 10. I have not detected predictability.
 
```{r}
#import viewership dataset
viewership <- read_csv("Datasets/viewers.csv") %>% 
              janitor::clean_names() %>% 
              pivot_longer(
              cols = series_1:series_10,
              names_to = "series",
              values_to = "viewership",
              names_prefix = "series_")%>%
              mutate(
              series = as.numeric(series),
              episode = as.numeric(episode))
  
knitr::kable(head(viewership, 10), caption = "First 10 rows")
```




