<<<<<<< HEAD
---
title: "P8105-HW2-lj2607"
author: "Li Jiang"
date: "2024-10-01"
output: github_document
---

#Problem 1

```{r}
#load the required packages
library(tidyverse)
library(readxl)
```

```{r}
#clean the data and retain the variable as required 
trans_ent = 
  read_csv(
    "Datasets/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) |> 
  janitor::clean_names() |> 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

#visually check the dataset
view(trans_ent)
```

```{r}
#check the size of dataset

nrow(trans_ent)
ncol(trans_ent)
```

Description of the dataset nyc_transit_clean:

I currently cleaned the original dataset and retained line, station_name, station_latitude, station_longitude, route, entry, exit_only, vending, entrance_type, ada. I also converted entry to logical variable. The sesulting dataset has `r ncol(trans_ent)` columns and `r nrow(trans_ent)` rows. The dimenison of the resulting data set is : `r dim(trans_ent)`. The dataset contain variables : `r names(trans_ent)`. As it stands, these data are not "tidy": route number should be a variable, as should route.

```{r}
#Number of distinct station
trans_ent |> 
  select(station_name, line) |> 
  distinct()
```
There are 465 distinct station.

```{r}
#Number of distinct station that are ADA compliant
trans_ent |> 
  filter(ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```
There are 84 ADA compliant station.

```{r}
#compute the proportion of station entrances / exits without vending allow entrance
trans_ent |> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```
The proportion of station entrances / exits without vending allow entrance is 0.3770492


```{r}
#identify stations that serve the A train
trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()
```

There are 60 distinct stations servcing the A train.

```{r}
#assess how many of these are ADA compliant
trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A", ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```

17 of these are ADA compliant


#Problem 2

```{r}
# load the package

library(tidyverse)
library(readxl)

# Import Mr.Trash Wheel dataset, specifying the sheet and range

mr_trash_data <- read_excel("Datasets/202409 Trash Wheel Collection Data.xlsx",
                             sheet = 'Mr. Trash Wheel',
                             range = "A2:N653",
                             na = c("NA", "", ".")) %>% 
                   janitor::clean_names() %>% # use reasonable variable names
                   drop_na(dumpster) %>%      # drop the rows without specific dumpster                                                           information
                   mutate(sports_balls = as.integer(round(sports_balls,digits = 0)),
                                              # convert sport ball to integer variable
                          trash_wheel = "mr_trash", # creating additional variable recording                                                            trash_wheel           
                          year = as.numeric(year)) # convert year to numerical variable
                                             

# visually inspect dataset
view(mr_trash_data)
```

```{r}
#Import Professor.Trash Wheel dataset, specifying the sheet and range

prof_trash_data <- read_excel("Datasets/202409 Trash Wheel Collection Data.xlsx",
                              sheet = 'Professor Trash Wheel',
                              range = "A2:M120",
                              na = c("NA", "", ".")) %>% 
                   janitor::clean_names() %>% # use reasonable variable names
                   drop_na(dumpster) %>%      # drop the rows without specific dumpster                                                           information
                   mutate(trash_wheel = "prof_trash") 
                                              # creating additional variable recording trash_wheel        

# visually inspect dataset
view(prof_trash_data)

#Total weight of trash collected by professor trash wheel

sum(prof_trash_data$weight_tons, na.rm = TRUE)
```

```{r}
#Import Gwynnda Trash Wheel dataset, specifying the sheet and range

gwynnda_trash_data <- read_excel("Datasets/202409 Trash Wheel Collection Data.xlsx",
                              sheet = 'Gwynnda Trash Wheel',
                              range = "A2:L265",
                              na = c("NA", "", ".")) %>% 
                   janitor::clean_names() %>% # use reasonable variable names
                   drop_na(dumpster) %>%      # drop the rows without specific dumpster                                                           information
                   mutate(trash_wheel = "gwynnda_trash") 
                                              # creating additional variable recording trash_wheel        

# visually inspect dataset
view(gwynnda_trash_data)

# total number of cigarette butts collected by Gwynnda in June of 2022
total <- gwynnda_trash_data %>%
         filter(month == "June", year == "2022") %>%
         summarise(total_cigarette_butts = sum(cigarette_butts, na.rm = TRUE))
total
```

```{r}
#Combine three sheets to produce a single tidy sheet
trash_wheel_data = 
  bind_rows(mr_trash_data, prof_trash_data, gwynnda_trash_data) %>% 
  janitor::clean_names()

# visually inspect dataset
view(trash_wheel_data)
```

Trash_wheel_data is composed of three datasets: mr_trash_data, prof_trash_data, gwynnda_trash_data. There are `r nrow(trash_wheel_data)` rows and `r ncol(trash_wheel_data)` in datase and key variable include `r names(trash_wheel_data)`. The number of observation in datset is `r nrow(trash_wheel_data)`. The total weight of trash collected by professor trash wheel is 246.74 tons. The total number of cigarette butts collected by Gwynnda in June of 2022 is 18120.

